---
title: "Reproducible Research Week 2 Project 1"
author: "Renga Pandurengan"
date: "6/30/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Loading required packages
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(Hmisc)
```

1. Loading & Manipulating data

```{r}
data <- read.csv("C:/Users/rkrishnaramanujam/Desktop/Data_Science/ReproducibleResearch/Week_2/Project_1/repdata_data_activity/activity.csv")
data$date_new <- as.Date(data$date) 

```

2. Creating Histogram of the total number of steps taken each day

```{r}
stepsPerDay <- tapply(data$steps, data$date, sum, na.rm=TRUE)
qplot(stepsPerDay, xlab='Total steps per day', ylab='Frequency ')
```

3. Mean and median number of steps taken each day
 
```{r}
 MeanstepsByDay <- mean(stepsPerDay)
MedianstepsByDay <- median(stepsPerDay)
```

4. Time series plot of the average number of steps taken

```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE)


ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
  geom_line() +
  xlab("Interval (5 minute)") +
  ylab("average number of steps taken") 
```

5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
maxSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMaxSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[maxSteps,'interval'])
```

6. Code to describe and show a strategy for imputing missing data


```{r}
#Number of missing values
NoMissingValues <- sum(is.na(data$steps))
NoMissingValues
#Strategy for imputing missing values
dataImputed <- data
dataImputed$steps <- impute(data$steps, fun=mean)
```

7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
# Create the histogram with imputed data.
qplot(dataImputed$steps, xlab='Total steps per day', ylab='Frequency ')
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
dataImputed$dateType <-  ifelse(as.POSIXlt(dataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
MeandataImputed <- aggregate(steps ~ interval + dateType, data=dataImputed, mean)

# Panel plot 

ggplot(MeandataImputed, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(dateType ~ .) +
  xlab("Interval") + 
  ylab("average number of steps")
```

9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

